---
title: "The 28th AHR Tables & Charts"
format: 
  html:
    toc: true
    toc-floating: true
    toc-title: "Contents"
    toc-depth: 2
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
editor: visual
---

```{r}
#| echo: false
library(readxl)
library(tidyr)
library(ggplot2)
library(stringr)
library(dplyr)
library(plotly)
library(maps)
library(sf)
library(DT)
library(janitor)
```

```{r}
AHR_data <- readxl::read_excel("data/AHR_data 28th_summing_components_columns.xlsx", sheet = 1)
score_rankings <- readxl::read_excel("data/AHR_data 28th_summing_components_columns.xlsx", sheet = 2)
disbursement_data <- readxl::read_excel("data/AHR_data 28th_summing_components_columns.xlsx", sheet = 3)
```

# TABLE 1: OVERALL HIGHWAY PERFORMANCE RANKINGS

```{r}
table1_2 <- score_rankings %>% select(state, overall_score_rank) %>% 
  arrange(overall_score_rank) %>% 
  datatable(option = list(lengtheMenu = c(10),
                          pageLength = 10))
```

# TABLE 2: OVERALL HIGHWAY PERFORMANCE RANKINGS

IN ALPHABETICAL ORDER

```{r}
score_rankings %>% select(state, overall_score_rank) %>% 
  arrange(state) %>% 
  datatable(option = list(lengtheMenu = c(10),
                          pageLength = 10))
```

# TABLE 3: HIGHWAY PERFORMANCE RANKINGS BY CATEGORY

```{r}
table3 <- score_rankings %>% 
    select(state, 
           overall_score_rank,
           #Disbursement
           capital_disbursement_perlm_score_rank, 
           maintenance_disbursement_perlm_score, 
           admin_disbursement_perlm_score_rank, 
           other_disbursement_perlm_score_rank,
           
           #Pavement roughness
           rural_interstate_poor_percent_score_rank,
           urban_interstate_poor_percent_score_rank,
           
           rural_OPA_poor_percent_score_rank,
           urban_OPA_poor_percent_score_rank,
           
           state_avg_congestion_hours_score_rank,
           
           poor_bridges_percent_score_rank,
           
           #fatality
           rural_fatalities_per_100m_VMT_score_rank,
           urban_fatalities_per_100m_VMT_score_rank,
           other_fatalities_per_100m_VMT_score_rank
           ) %>% datatable()


```

# TABLE 4: OVERALL HIGHWAY PERFORMANCE RANKING TRENDS

```{r}
table4 <- read.csv("data/Score_Rankings_in27threport.csv") %>% 
  clean_names() %>% 
  select(1:4) %>% 
  left_join(score_rankings %>% select(state, overall_score_rank)) %>% 
  rename(x2021 = overall_score_rank) %>% 
  mutate(change20_21 = (x2020 - x2021),
         change19_21 = (x2019 - x2021),
         change18_21 = (x2018 - x2021)) %>% 
  datatable(option = list(lengtheMenu = c(10),
                          pageLength = 10))

```

# TABLE 5: STATE-CONTROLLED HIGHWAY MILES, 2020

```{r}
table5 <- AHR_data %>% 
  arrange(desc(state_tot_lane_miles)) %>% 
  mutate(`2021 Size` = ifelse(row_number() == 1, "-", row_number() - 1)) %>% 
  select(`2021 Size`, state, state_tot_lane_miles) %>% 
  datatable(rownames = FALSE) %>% 
  formatStyle("state_tot_lane_miles",'text-align' = 'right') %>% 
  formatRound("state_tot_lane_miles", digits = 0)
  
```

# TABLE 6: STATE-CONTROLLED HIGHWAY MILEAGE BY SYSTEM WIDTH, 2020

Note: Missing "Centerline Mileage"

```{r}

table6 <- AHR_data %>% 
  arrange(desc(state_tot_lane_miles)) %>% 
  mutate(`2021 Size` = ifelse(row_number() == 1, "-", row_number() - 1)) %>% 
  select(`2021 Size`, state, SHA_ratio, state_tot_lane_miles) %>% 
  arrange(desc(SHA_ratio)) %>% 
  
  datatable(rownames = FALSE) %>% 
  formatStyle(3:4,'text-align' = 'right') %>% 
  formatRound("SHA_ratio", digit = 2) %>% 
  formatRound("state_tot_lane_miles", digit = 0)
```

# TABLE 7: CAPITAL AND BRIDGE DISBURSEMENTS

```{r}
table_disbursement <- function (disbursement_cat) {disbursement_data %>% 
  filter(str_detect(key_metrics, disbursement_cat)) %>% 
  select(-key_metrics) %>% 
  mutate(adjusted_ratio = value/exp_value) %>% 
  arrange(value) %>% 
  mutate(`2021 Rank` = row_number()) %>% 
  datatable(rownames = FALSE) %>% 
  formatStyle(2:4,'text-align' = 'right') %>% 
  formatRound(2:3, digit = 0) %>% 
  formatRound(4, digit = 2) 
}

table7 <- table_disbursement("capital")
```

# TABLE 8: MAINTENANCE DISBURSEMENTS

```{r}
table8 <-table_disbursement("maintenance")
```

# TABLE 9: ADMINISTRATIVE DISBURSEMENTS

```{r}
table9 <-table_disbursement("admin")
```

# TABLE 10: OTHER DISBURSEMENTS

```{r}
table10 <- table_disbursement("other")
```

# TABLE 11: PERCENT RURAL INTERSTATE MILEAGE IN POOR CONDITION

NOTE: missing Weighted Average

```{r}
table_cat <- function(cat) {
  AHR_data %>% select(state, {{cat}}) %>% 
  arrange(is.na({{cat}}), {{cat}}) %>% 
  mutate(`2021 Rank` = row_number()) %>% 
  datatable(rownames = FALSE) %>% 
  formatStyle(2,'text-align' = 'right') %>% 
  formatRound(2, digit = 2)
}

table11 <-table_cat(rural_interstate_poor_percent)
```

# TABLE 12: PERCENT URBAN INTERSTATE MILEAGE IN POOR CONDITION

```{r}
table12 <-table_cat(urban_interstate_poor_percent)
```

# TABLE 13: PERCENT RURAL OTHER PRINCIPAL ARTERIAL MILEAGE IN POOR CONDITION

```{r}
table13 <-table_cat(rural_OPA_poor_percent)
```

# TABLE 14: PERCENT URBAN OTHER PRINCIPAL ARTERIAL MILEAGE IN POOR CONDITION

```{r}
table14 <-table_cat(urban_OPA_poor_percent)
```

# TABLE 15: ANNUAL PEAK HOURS SPENT IN CONGESTION PER AUTO COMMUTER

```{r}
table15 <-table_cat(state_avg_congestion_hours)
```

# TABLE 16: PERCENT STRUCTURALLY DEFICIENT BRIDGES

```{r}
table16 <-table_cat(poor_bridges_percent)
```

# TABLE 17: FATALITY RATE PER 100 MILLION RURAL VEHICLE-MILES

```{r}
table17 <-table_cat(rural_fatalities_per_100m_VMT)
```

# TABLE 18: FATALITY RATE PER 100 MILLION URBAN VEHICLEMILES

```{r}
table18 <-table_cat(urban_fatalities_per_100m_VMT)
```

# TABLE 19: FATALITY RATE PER 100 MILLION OTHER VEHICLE-MILES

```{r}
table19 <-table_cat(other_fatalities_per_100m_VMT)
```
